---
title: "Comparison of trade data sources"
author: "Aleksandr Matrunich"
date: "June 12, 2015"
output: 
  html_document:
    toc: true
---
Last update: `r Sys.time()`


# Introduction

We want to compare ComTrade data from TF_SOURCE and SWS


# Getting data and calculation

```{r, echo = F, warning=F}
knitr::opts_knit$set(width = 100)
library(fclhs)
suppressPackageStartupMessages(library(dplyr))

# Class for accessing DB with OJDBC
.ojdbcclasspath <- "/home/sas/dstrbs/ojdbc14.jar"

source("/home/sas/sws/sws_r_api/r_modules/trade_prevalidation/vignettes/setupconnection.R")

```


Getting data from TF_SOURCE*

```{r, warning=F}
tfsource <- bind_rows(
  cbind(gettfsource(reporter = c(231, 105, 79), year = 2010), year = 2010),
  cbind(gettfsource(reporter = c(231, 105, 79), year = 2011), year = 2011))

head(tfsource, 10)
```

## What is missing value in TF_SOURCE

How to calculate missing values? Is there NAs?

```{r}
tfsource %>% 
  summarize(qty_na  = sum(is.na(qty)),
            qty2_na = sum(is.na(qty2)),
            qty_0   = sum(qty == 0),
            qty2_0  = sum(qty2 == 0))
```

Only zeros and no NAs. So we'll use zeros as missing values.

## How many digits in HS-codes from TF_SOURCE?

```{r}
tfsource %>% 
  select(hs) %>% 
  mutate(hsdigits = stringr::str_length(hs)) %>% 
  group_by(hsdigits) %>% 
  summarize(flowstotal = n())
```

OK, so to compare with SWS we need to aggregate down to 6 digits.

## Trimming to 6-digits HS codes

```{r}
tfsource <- tfsource %>% 
  mutate(hs = stringr::str_extract(hs, "^.{6}")) %>% 
  group_by(year, reporter, partner, flow, hs) %>% 
  summarize(qty = sum(qty),
            qty2 = sum(qty2),
            value = sum(value))
```


## Are there other than agricultural items?

```{r}
any(!(unique(tfsource$hs) %in% getAgriHSCodes()))
```
OK, they are. So we need them to be filter out.


## Removing non-agricultural HS-items

```{r}
tfsource <- data.frame(hs = getAgriHSCodes(), stringsAsFactors = F) %>% 
  inner_join(tfsource, by = "hs")
any(!(unique(tfsource$hs) %in% getAgriHSCodes()))
```

## Aggregation

```{r}
tfsource_agg <- tfsource %>% 
  group_by(year, reporter) %>% 
  summarize(totalflows  = n(),
            noqty       = sum(qty == 0 & qty2 == 0 & value > 0),
            noqty_prcnt = round(100 * noqty / totalflows, 2),
            novl        = sum(qty > 0 & qty2 > 0 & value == 0),
            novl_prcnt  = round(100 * novl / totalflows, 2))
```

And for comparison similar data from SWS.

```{r}
swsdata <- readRDS("/home/sas/sws/sws_r_api/USGEIL_agri.rds") %>% 
  humanizeComtrade() %>% 
  select(-unit) %>%
  reshape2::dcast(... ~ group) %>%  
  group_by(reporter) %>% 
  summarize(
    totalflows = n(),
    noquant = sum(((weight == 0 | is.na(weight)) &
                     (volume == 0 | is.na(volume)) &
                     (head == 0 | is.na(head))) &
                    cost > 0, 
                  na.rm = T),
    noquantprcnt   = round(100 * noquant / totalflows, 2),
    novalue = sum((weight > 0 | volume > 0 | head > 0) & 
                    (is.na(cost) | cost == 0), na.rm = T),
    novalueprcnt   = round(100 * novalue / totalflows, 2))
```
# --> Final tables for comparison <--
SWS data only for 2011 year. I've added 2010 year in TF_SOURCE table to be sure that the absence of Germany's data is year-specific.
```{r final_tables}
swsdata

tfsource_agg
```
We observe similar amounts of trade flows for Israel (105) and USA (231). Amount of trade flow with no quantity differs much for Israel.

# Investigation of absence of Germany data in TF_SOURCE

In TFSOURCE there is no data for Germany (79) from 2011! Let's check other sources on small example (Bovine).

Data from UN ComTrade [public API](http://comtrade.un.org/data/).

```{r, warning=F, message=F}
unct <- getunct(area = paste(unctarea(79), unctarea(231), sep = ","), 
                      year = "2011, 2010",
                      partner = unctarea(fclhs:::findfaoareacode("Canada")),
                      code = "010210",
                      desc = T,
                      compact = F)
unct
```

There is data for Germany only for 2011 year and its flow has flag 2. I didn't find description of the flags.

Similar data from TF_SOURCE.

```{r, message=F}
tfsource %>% 
  filter(reporter %in% c(231, 79),
         stringr::str_detect(hs, "^010210"),
         partner == fclhs:::findfaoareacode("Canada"))
```

It has 10 digits in HS code. Let's aggregate to 6 digits.

```{r, message=F}
tfsource %>% 
  filter(reporter %in% c(231, 79),
         stringr::str_detect(hs, "^010210"),
         partner == fclhs:::findfaoareacode("Canada")) %>% 
  mutate(hs = stringr::str_extract(hs, "^.{6}")) %>% 
  group_by(year, reporter, partner, flow, hs) %>% 
  summarize(qty = sum(qty),
            qty2 = sum(qty2),
            value = sum(value))
# And public API Comtrade
unct %>% select(yr, flow = rgDesc, reporter = rtTitle, partner = ptTitle, hs = cmdCode, qtDesc,
                qty = TradeQuantity, kg = NetWeight, value = TradeValue, flag = estCode)
```

```{r}
getComtradeData(reporter = c(842, 276), partner = 124, year = 2010:2011, item = "010210", 
                element  = selectElems(unit %in% c("kg", "US$", "t", "1000 t", "head", "1000 head"))) %>% 
  humanizeComtrade() %>% 
  group_by(year, reporter, dir, unit)
```

