---
title: "M49 country codes in Tariffline"
author: "Aleksandr Matrunich"
date: "July 13, 2015"
output: html_document
---
Last update: `r Sys.time()`

```{r options, echo = F, warning=F}
options(width = 120)
suppressPackageStartupMessages(library(dplyr))

source(file.path(Sys.getenv("HOME"), "r_adhoc", "trade_prevalid_testing", "setupconnection.R"))

subdir <- "OrangeBook"
sourcedir <- "tradeR"

if(length(lapply(dir(file.path(Sys.getenv("HOME"), "r_adhoc", "privateFAO", subdir, sourcedir), 
                     full.names = T), 
                 source)) == 0) stop("Files for sourcing not found")

trade_src <- src_postgres("sws_data", "localhost", 5432, "trade", .pwd, 
                          options = "-c search_path=ess")

cttl <- tbl(trade_src, sql("select * from tariffline"))
```

Codes of reporting countries which are absent among partner countries.

```{r absent_reporters}

# http://explainextended.com/2009/09/16/not-in-vs-not-exists-vs-left-join-is-null-postgresql/

  absent_reps <-  tbl(trade_src, sql("select distinct reporter
from tariffline tr
where not exists (select 1
                   from tariffline tp
                  where tr.reporter = tp.partner)")) %>% collect() %>% 
    '[['("reporter")
```

```{r absent_partners}
# absent_prts <- tbl(trade_src, sql("select distinct partner from tariffline 
#                    left join 
#                    (select  distinct reporter from tariffline) as tbl1
#                    on tariffline.partner = tbl1.reporter 
#                    where tbl1.reporter is null ")) %>% collect() %>% 
#   '[['("partner")
```

```{r reporter_codes_in_sws}
getCountryName(absent_reps, drop = F)
# getCountryName(absent_prts, drop = F) %>% arrange(code) %>% as.data.frame()
```

## Possible conflicts between official M49 codes scheme and Comtrade one

Codes from official M49[^officialsource] which are absent in Comtrade M49[^comtradesource].

[^officialsource]: http://unstats.un.org/unsd/methods/m49/m49alpha.htm Countries or areas, codes and abbreviations 

[^comtradesource]: http://comtrade.un.org/data/cache/reporterAreas.json The UN Comtrade data extraction API


```{r official_m49_absent_in_ct}
officialm49 <- getOfficialM49()
comtradem49 <- getComtradeM49()

officialm49[!is.element(as.integer(officialm49$code),
                        as.integer(comtradem49$code)), ] %>% 
  arrange(code)
```

Codes from Comtrade M49 which are absent in Official M49.
```{r comtrade_m49}
comtradem49[!is.element(as.integer(comtradem49$code), 
                        as.integer(officialm49$code)), ] %>% 
  arrange(code)

```

It would be useful to check matching of codes based on strings comparing.

## Coverage of codes from Tariffline by official and Comtrade M49
### Reporters not found in Comtrade scheme
```{r}
reps <- tbl(trade_src, sql("select distinct reporter from tariffline")) %>%
  collect()

reps %>% 
  mutate(reporter = as.integer(reporter)) %>% 
  anti_join(getComtradeM49() %>% 
              mutate(code = as.integer(code)),
            by = c("reporter" = "code"))
  
```
No such partners. So column with partners is standardized in Tariffline.

### Partners not found in official scheme
```{r}

prts <- tbl(trade_src, sql("select distinct partner from tariffline")) %>%
  collect()

prts_not_official <- prts %>% 
  mutate(partner = as.integer(partner)) %>% 
  anti_join(getOfficialM49() %>% 
              mutate(code = as.integer(code)),
            by = c("partner" = "code")) %>% 
  unlist() %>% unname() %>% sort()

prts_not_official 
```

### Partners not found in Comtrade scheme
```{r partners_not_comtrade}
prts_not_comtrade <- prts %>% 
  mutate(partner = as.integer(partner)) %>% 
  anti_join(getComtradeM49() %>% 
              mutate(code = as.integer(code)),
            by = c("partner" = "code")) %>% 
unlist() %>% unname() %>% sort()

prts_not_comtrade 
```

Column with reporters is not standardized in Tariffline, but official M49 look better.

### Partners not found in SWS

```{r partners_not_sws}
prts_not_sws <- getCountryName(prts$partner, drop = F) %>% 
  filter(is.na(description)) %>% 
  select(code) %>% 
  arrange(code) %>% 
  unlist() %>% unname()

prts_not_sws
```

Wow! SWS is a champion!

### Mysterious partners
```{r mysterious_reporters}
intersect(prts_not_comtrade, 
          intersect(prts_not_sws, 
                    prts_not_official))
```

## TODO

* Check possible conflicts between similar codes from different sources
* How big is data with mysterious partners?
