---
title: 'Intersection of MDB and Tariffline sets of HS-codes'
author: "Aleksandr Matrunich"
date: "July 6, 2015"
output: 
  html_document:
    toc: true
---
Last update: `r Sys.time()`

```{r options, echo = F, warning=F}
options(width = 120)
.pwd <- "KlO%0cTEm"
suppressWarnings(library(fclhs))
suppressPackageStartupMessages(library(dplyr))

# Class for accessing DB with OJDBC
.ojdbcclasspath <- file.path(Sys.getenv("HOME"), "dstrbs", "ojdbc14.jar")

source(file.path(Sys.getenv("HOME"), "r_adhoc", "trade_prevalid_testing", "setupconnection.R"))

trade_src <- src_postgres("sws_data", "localhost", 5432, "trade", .pwd, 
                          options = "-c search_path=ess")

cttl <- tbl(trade_src, sql("select * from tariffline"))
```

```{r read_mdbdata}

mdbdata <- data.table::fread(file.path(Sys.getenv("HOME"), "r_adhoc", "mdb_read", "hs_fcl_full.csv"),
                              data.table = F)


mdbdata$validyear[mdbdata$validyear == ""] <- "0"
mdbdata$validyear <- as.integer(mdbdata$validyear)

head(mdbdata)
```

Proportion of records where there is only one code and not a range of codes: `r 
mdbdata %>% 
  summarize(singlecode = paste0(round(100 * sum(fromcode == tocode) / n(), 1), "%")) %>% 
  unlist() %>% unname()`

```{r tariff_ex}
tbl(trade_src, sql("select * from tariffline limit 10"))

```

```{r us2011_collect}

agriItems <- paste0("'", paste(getAgriHSCodes(), collapse = "','"), "'")

tl_us2011 <- tbl(trade_src, sql(paste0(
"select * from (select year, reporter, partner, flow, hs, substring(hs from 1 for 6) AS hs6  
from tariffline
where reporter = '842' and year = '2011') as tbl1 
where hs6 in(", agriItems, ") "))) %>% 
  select(-hs6) %>% collect()

mdb_us2011 <- mdbdata %>% 
  select(-area, -iso2, -country, -cpc) %>% 
  filter(fao == faoarea(842), 
         validyear %in% c("0", "2011"))
```

Proportion of records where there is only one code and not a range of codes: `r 
mdb_us2011 %>% 
  summarize(singlecode = paste0(round(100 * sum(fromcode == tocode) / n(), 1), "%")) %>% 
  unlist() %>% unname()`
  
  
OK, so for US 2011 there is no ranges of HS codes so we can postpone problem of matching HS-code to range from mdb data.

Degree of matching HS-codes between MDB base and TariffLine.

```{r simple_match}

tl_us2011 %>% 
  filter(flow == "1") %>% 
  select(hs) %>% 
  distinct() %>%  
  anti_join(mdb_us2011 %>% 
              filter(flow == "Import") %>% 
              select(hs = fromcode, fcl) %>% 
              distinct(), 
            by = "hs") %>% 
  nrow()
```

