---
title: 'Tariff line data: first impressions'
author: "Aleksandr Matrunich"
date: "June 25, 2015"
output: 
  html_document:
    toc: true
---
Last update: `r Sys.time()`

```{r options, echo = F, warning=F}
options(width = 120)
.pwd <- "KlO%0cTEm"
library(fclhs)
suppressPackageStartupMessages(library(dplyr))

# Class for accessing DB with OJDBC
.ojdbcclasspath <- "/home/sas/dstrbs/ojdbc14.jar"

source("/home/sas/sws/sws_r_api/r_modules/trade_prevalidation/vignettes/setupconnection.R")

tlsql <- function(select = NULL, where = NULL) {
  
  if(!is.null(select)) select <- list(select)
  
  faosws::GetTableData(schemaName = "ess",
                       tableName = "ct_tariffline_adhoc", 
                       whereClause = where, 
                       selectColumns = select)
  
  
}

trade_src <- src_postgres("sws_data", "localhost", 5432, "trade", .pwd, 
                          options = "-c search_path=ess")

cttl <- tbl(trade_src, "ct_tariffline_adhoc")

```

## Introduction

We got tariff line data from UNSD. It's stored in ad hoc table `ess.ct_tariffline_adhoc` at QA SWS server. Currently it's possible to access the data only from R.

## Structure of Tariff Line table

### Size of dataset

For all countries and years.

```{r nrow}
cttl %>% summarize(nrows = n())
```

### Example of data
```{r structure}
cttl


tbl(trade_src, sql(
"select * 
from ess.ct_tariffline_adhoc 
where rep in ('376', '842', '276') order by random() limit 20")) %>%
  collect() %>% as.data.frame()

```
### Range of years

For all countries

```{r years_range}
cttl %>% 
  select(tyear) %>% 
  group_by(tyear) %>% 
  summarize(nrows = n()) %>% 
  arrange(tyear) %>% collect()
```

### Type of classification used by reporter

For all countires.

```{r classific_type}
tlsql(select = "hsrep, count(hsrep)", where = "group by hsrep order by hsrep") %>%  
  mutate(hsrep = factor(hsrep, levels = c("H0", "H1", "H2", "H3", "H4", "S3"),
                        labels = c("HS1992", "HS1996", "HS2002", 
                                   "HS2007", "HS2012", "SITCr3")))
```

### Lenght of hs codes

#### Full data set

```{r hs_length}
tbl(trade_src, sql("
select hs_length, count(hs_length) as countn 
from (select char_length(comm) as hs_length 
from ess.ct_tariffline_adhoc) as tbl1 
group by hs_length order by hs_length
                   ")) %>% collect()
```

#### Countries of the question

```{r hs_length_subset}
tbl(trade_src, sql("
select hs_length, count(hs_length) as countn 
from (select char_length(comm) as hs_length 
from ess.ct_tariffline_adhoc where rep in ('376', '842', '276')) as tbl1 
group by hs_length order by hs_length
                   ")) %>% collect()
```

### Reported flows

For all countires.

```{r flows}
tlsql(select = "flow, count(flow)", where = "group by flow order by flow") 
```

### Reported units

For all countires.

```{r units}
tlsql(select = "qunit, count(qunit)", where = "group by qunit order by qunit") 
```
See description of units in Annex I of [Quantity and Weight Data in UN Comtrade](http://unstats.un.org/unsd/tradekb/Knowledgebase/Quantity-and-Weight-Data-in-UN-Comtrade).

### Are there estimations?

For all countires.

```{r est}
# tlsql(select = "est, count(est)", where = "group by est order by est") 
tlsql(select = "sum( case when est != '0' then 1 else 0 end) as estimations") 
```
There is no estimated values.

### What does contain ht column?

```{r ht_column}
tlsql(select = "ht, count(ht)", where = "group by ht") 
```

There is only zero in column ht.


## What is missing value in Tariff Line data

It could be zero or NULL. We already saw NA weight and qty in random sample.

For all countries.

```{r missings}

miss <- tbl(trade_src, sql("select sum (case when weight is null then 1 else 0 end) as weight_null,
      sum (case when qty is null then 1 else 0 end) as qty_null,
      sum (case when tvalue is null then 1 else 0 end) as tvalue_null,
      sum (case when weight = 0 then 1 else 0 end) as weight_0,
      sum (case when qty = 0 then 1 else 0 end) as qty_0,
      sum (case when tvalue = 0 then 1 else 0 end) as tvalue_0 from ess.ct_tariffline_adhoc ")) %>% 
  collect() %>% as.integer()

dim(miss) <- c(3, 2)
colnames(miss) <- c("NA", "zero")
rownames(miss) <- c("weight", "qty", "tvalue")
miss

```

We have null (in R they are converted into NA) cases for weight and qty and zero cases for value.

## Conflicts between units

Are there cases when for the same trade flow we have several records with different units (note from Giorgio)?

```{r units_conflicts}
agriItems <- paste0("'", paste(getAgriHSCodes(), collapse = "','"), "'")

cttlagri <- tbl(trade_src, sql(paste0(
"select * 
from (select rep, tyear, flow, comm, substring(comm from 1 for 6) as hs6, 
prt, weight, qty, qunit, tvalue 
from ess.ct_tariffline_adhoc 
where rep in ('376', '842', '276') 
and tyear = '2011') as tbl1 
where hs6 in(", agriItems, ") "))) %>%
  select(-hs6)

cttlagri %>%
  group_by(tyear, rep, prt, flow, comm)  %>% 
  summarize(nrows = n()) %>% 
  ungroup() %>% 
  filter(nrows > 1)
  
```

There are now such cases for the pilot countries, but they exist in other countries. For example:

```
   tyear rep prt flow       comm nrows
1   2011 328 780    1 2106909000     5
2   2011 499 899    1     220290     4
3   2011 328 531    2 0709909000     4
4   2011 328 531    2 1904200000     4
5   2011 328 780    1 2009112000     4
6   2011 328 780    1 2203001000     4
7   2011 328 780    1 2202909000     4
8   2011 328 780    1 2202101000     4
9   2011 328 780    2 2208401000     4
10  2011 328 840    1 0402999000     4
```

```{r units_conflicts_case}

cttl %>% 
  select(-est, -ht) %>%
  filter(tyear == "2011",
         rep == "328",
         prt == "780",
         comm == "2106909000")
```

# Missing values

```{r missing_values}
cttl_missing <- cttlagri %>% 
  select(rep, tyear, flow, comm, prt, weight, qty, qunit, tvalue) %>% 
  filter(rep %in% c('376', '842', '276'),
         tyear == "2011") %>% 
  collect() %>% 
  group_by(tyear, rep) %>% 
  summarize(totalflows  = n(),
            noqty       = sum(is.na(qty)  & is.na(weight) & tvalue > 0),
            noqty_prcnt = round(100 * noqty / totalflows, 2),
            novl        = sum((!is.na(qty) | !is.na(weight)) & tvalue == 0),
            novl_prcnt  = round(100 * novl / totalflows, 2)) %>% 
  mutate(reporter = getCountryName(rep)) %>% 
  select(reporter, totalflows, noqty, noqty_prcnt, novl, novl_prcnt)

tfsource <- gettfsource(reporter = c(231, 105, 79), year = 2011) %>% 
  mutate(hs6 = stringr::str_extract(hs, "^.{6}"))

tfsource <- data.frame(hs = getAgriHSCodes(), stringsAsFactors = F) %>% 
  inner_join(tfsource, by = c("hs" = "hs6")) %>% 
  select(-hs) %>% 
  rename(hs = hs.y) %>% 
  mutate(reporter = getCountryName(unctarea(reporter)))
  

tfsource %>% 
  group_by(reporter) %>% 
  summarize(totalflows  = n(),
            noqty       = sum(qty == 0 & qty2 == 0 & value > 0),
            noqty_prcnt = round(100 * noqty / totalflows, 2),
            novl        = sum((qty > 0 | qty2 > 0) & value == 0),
            novl_prcnt  = round(100 * novl / totalflows, 2)) 

cttl_missing
```

### Why do sizes of TariffLine and TF_SOURCE differ?

#### Differencies in sets of unique HS codes

```{r sets_of_codes}
tfsource_uniq_hs <- tfsource %>% 
  select(hs) %>% 
  distinct() %>% 
  collect()

tariffline_uniq_hs <- cttlagri %>% 
  select(rep, tyear, comm) %>% 
  filter(rep %in% c('376', '842', '276'),
         tyear == "2011") %>% 
  select(comm) %>% 
  distinct() %>% 
  select(hs = comm) %>%
  collect()

  data.frame(source = c("tfsource", "tariffline"),
             uniq_hs = c(nrow(tfsource_uniq_hs),
                         nrow(tariffline_uniq_hs)))
  
  # HS codes which are absent in TF_SOURCE

  tariffline_uniq_hs %>% 
    anti_join(tfsource_uniq_hs, by = "hs") %>% 
    sample_n(50) %>% 
    unlist() %>% unname() %>% as.factor() 
  
  # HS codes which are absent in Tariff Line
  
  tfsource_uniq_hs %>% 
    anti_join(tariffline_uniq_hs, by = "hs") %>% 
    sample_n(50) %>% 
    unlist() %>% unname() %>% as.factor() 
  
  # Trimmed to 6 digits HS codes which are absent in TF_SOURCE
  
  tariffline_uniq_hs %>% 
    mutate(hs = stringr::str_extract(hs, "^.{6}")) %>% 
    anti_join(tfsource_uniq_hs %>% 
                mutate(hs = stringr::str_extract(hs, "^.{6}")),
              by = "hs") 
  
  # Trimmed to 6 digits HS codes which are absent in Tariff Line
  
  tfsource_uniq_hs %>% 
    mutate(hs = stringr::str_extract(hs, "^.{6}")) %>% 
    anti_join(tariffline_uniq_hs %>% 
                mutate(hs = stringr::str_extract(hs, "^.{6}")),
              by = "hs") 
```



```{r close_conn, echo = F, results = "hide"}
RPostgreSQL::dbDisconnect(trade_src$con)
```
