---
title: "Trade2"
author: "Aleksandr Matrunich"
date: "August 8, 2015"
output: pdf_document
---

```{r set_options, echo = F}
suppressPackageStartupMessages(library(dplyr))

knitr::opts_chunk$set(list(echo = FALSE,
                           eval = TRUE,
                           cache = TRUE,
                           warning = FALSE,
                           message = FALSE,
                           fig.height = 10,
                           fig.width = 12))

hostname <- Sys.getenv("HOSTNAME")
if(!is.element(hostname, c("matrunichstation", "sasmobile2.sasdomain"))) 
  stop("Uknown workstation. Aborting.")

if(hostname == "sasmobile2.sasdomain") projects_dir <- "fao"

subdir <- "OrangeBook"
sourcedir <- "tradeR"

if(length(lapply(dir(file.path(Sys.getenv("HOME"), projects_dir, "privateFAO", subdir, sourcedir), 
                     full.names = T), 
                 source)) == 0) stop("Files for sourcing not found")

if(hostname == "sasmobile2.sasdomain") 
  orange <- readRDS(file.path(Sys.getenv("HOME"), 
                              projects_dir, 
                              "trade_prevalid_testing",
                              "tariffline_agri_orange.Rds"))

comtrade_areas <- getComtradeM49() %>% 
  mutate(
    common_name = countrycode::countrycode(name,
                                           "country.name",
                                           "country.name",
                                           warn = T))
official_areas <- getOfficialM49() %>% 
  mutate(
    # We take iso3 as source because it provides the best cover (only 2 NA)
    common_name = countrycode::countrycode(iso3,
                                             "iso3c",
                                             "country.name",
                                             warn = T))

orange %>% 
  left_join(getOfficialM49(),
            by= c("reporter" = "code"))
```

# Validation of trade data

At prevalidation step we are to make a decision should we accept data from a specific country for the further processing or not. A country could provide data of good quality for one part of commodities and inadequate level of quality for another part. We want to estimate quality differences between commodities of a country.

Quality of data is estimated by following indicators:

* Share per cent of missing quantities
* Share per cent of unit value outliers


### Self-trade

There are cases when a country reports itself as a partner to exports or imports. Such situations can occur due to mistakes or when an entrepôt exists.

In case of the US 2011 there are no cases of self-trade.

```{r examples_of_selftrade}
orange %>% 
  filter(reporter == partner)


```


## Detection of outliers

## Imputing of missing quantities and replacement of outliers