---
title: "Trade2"
author: "Aleksandr Matrunich"
date: "August 8, 2015"
output: pdf_document
---

```{r set_options, echo = F, warning = F}
suppressPackageStartupMessages(library(dplyr))
library(ggplot2)

knitr::opts_chunk$set(list(echo = FALSE,
                           eval = TRUE,
                           cache = TRUE,
                           warning = FALSE,
                           message = FALSE,
                           fig.height = 10,
                           fig.width = 12))

hostname <- Sys.getenv("HOSTNAME")
if(!is.element(hostname, c("matrunichstation", "sasmobile2.sasdomain"))) 
  stop("Uknown workstation. Aborting.")

if(hostname == "sasmobile2.sasdomain") projects_dir <- "fao"

subdir <- "OrangeBook"
sourcedir <- "tradeR"

if(length(lapply(dir(file.path(Sys.getenv("HOME"), projects_dir, "privateFAO", subdir, sourcedir), 
                     full.names = T), 
                 source)) == 0) stop("Files for sourcing not found")

if(hostname == "sasmobile2.sasdomain") 
  orange <- readRDS(file.path(Sys.getenv("HOME"), 
                              projects_dir, 
                              "trade_prevalid_testing",
                              "tariffline_agri_orange.Rds"))

orange <- orange %>% 
  left_join(getComtradeM49()  %>%
              mutate(
                reporter_name = countrycode::countrycode(name,
                                                       "country.name",
                                                       "country.name",
                                                       warn = F)) %>% 
              select(-name),
            by= c("reporter" = "code")) %>% 
  left_join(getOfficialM49() %>% 
              select(-name) %>% 
              mutate(
                # We take iso3 as source because it provides the best cover (only 2 NA)
                partner_name = countrycode::countrycode(iso3,
                                                        "iso3c",
                                                        "country.name",
                                                        warn = F)) %>% 
              select(-iso3),
            by = c("partner" = "code")) %>% 
  mutate(flow = factor(flow, levels = c(1, 2, 3, 4), 
                       labels = c("Import", "Export",
                                  "Re-Import", "Re-Export")))
```

# Initial validation of trade data

At prevalidation step we are to make a decision should we accept data from a specific country for the further processing or not. A country could provide data of good quality for one part of commodities and inadequate level of quality for another part. We want to estimate quality differences between commodities of a country.

Quality of data is estimated by following indicators:

* Share per cent of missing quantities
* Share per cent of unit value outliers


## Self-trade

There are cases when a country reports itself as a partner to exports or imports. Such situations can occur due to mistakes or when an entrepôt exists.

```{r examples_of_selftrade}
orange %>% 
  filter(reporter_name == partner_name) %>% 
  group_by(reporter_name, flow) %>% 
  summarize(ntradeflows = n()) %>% 
  ungroup() %>% 
  arrange(desc(ntradeflows)) %>% 
  rename(Reporter = reporter_name, 
         Flow = flow, 
         Total = ntradeflows) %>% 
  printTab(caption = "Self-trade of commodities from 2nd, 10th and 15th HS chapters in 2011")
```

## Missing quantity

We identify which reporters provide data of insufficient quality. Firstly for every reporter proportion of trade flows with missing quantity is calculated. 

```{r missing_quantity_props_per_country, cache = F, fig.width = 5, fig.height = 5, dpi = 200}
no_quant_threshold <- .02

orange <- orange %>% 
  mutate(no_quant = missingIndicator(weight, NA) &
           missingIndicator(qty, NA))  

missing_qty <- orange %>% 
  group_by(reporter_name) %>% 
  summarize(missing_prop = sum(no_quant) / n()) %>% 
  ungroup() %>% 
  filter(missing_prop > no_quant_threshold) %>% 
  arrange(desc(missing_prop))

missing_qty %>% 
  mutate(reporter_name = stringr::str_replace(
    reporter_name, ",.*$", "...")) %>% 
  ggplot(aes(missing_prop, reorder(reporter_name, missing_prop))) +
  geom_point() +
  scale_x_continuous("Trade flows with missing quantities",
                     labels = scales::percent) +
  scale_y_discrete("")
```

## Detection of outliers

We define outliers as observations located outside the range:

$$ [ Q_1 - k (Q_3 - Q_1 ) , Q_3 + k (Q_3 - Q_1 ) ] $$

where $Q_1$ and $Q_3$ are the lower and upper quartiles respectively, and $k$ is a nonnegative constant. In this paper we use $k$ equals 1.5. 

```{r prop_of_outliers_per_country, fig.width = 5, fig.height = 5, dpi = 300}
out_coef <- 1.5

orange <- orange %>% 
  mutate(
    qty_all = ifelse(!is.na(weight), weight, qty),
    uv = value / qty_all) %>% 
  select(-qty_all) %>% 
  group_by(reporter_name, hs) %>% 
  mutate(
    uv_reporter = median(uv, na.rm = T),
    out_range = uv %in% boxplot.stats(uv, coef = out_coef, do.conf = F)) %>% 
  ungroup()

orange %>%
  group_by(reporter_name) %>% 
  summarize(prop_of_outs = sum(out_range) / n()) %>%
  top_n(30, prop_of_outs) %>%
  ggplot(aes(prop_of_outs, reorder(reporter_name, prop_of_outs))) +
  geom_point() +
  scale_x_continuous("Trade flows with unit value outliers",
                     labels = scales::percent) +
  scale_y_discrete("")
```

## Missing quantites and outliers combined

```{r missings_outliers_together}
orange %>% 
  group_by(reporter_name) %>% 
  summarize(prop_of_miss = sum(no_quant) / n(),
            prop_of_outs = sum(out_range) / n()) %>% 
  filter(prop_of_miss > .001 & prop_of_outs > .001) %>% 
  ggplot(aes(x = prop_of_miss, 
             y = prop_of_outs,
             label = reporter_name)) +
  geom_text(size = 5) + 
  scale_x_log10(
    "Trade flows with missing quantity. Log10 scale",
    breaks = c(.002, .01, .1, .5, 1),
    labels = scales::percent) +
  scale_y_continuous(
    "Trade flows with UV-outliers",
    labels = scales::percent)
```



# Imputing of missing quantities and replacement of outliers

In data reported by USA for 2011 year in HS chapters 2, 10 and 15 there are `r 
orange %>% 
  filter(reporter_name == "United States" & no_quant) %>% 
  nrow` trade flows with missing quantity and `r 
orange %>% 
  filter(reporter_name == "United States" & out_range) %>% 
  nrow` trade flows with UV-outliers.
  
```{r us_outliers_example}
out_tab <- orange %>% 
  filter(reporter_name == "United States" & out_range & !is.na(partner_name)) %>% 
  select(Reporter = reporter_name, 
         Partner  = partner_name,
         Flow     = flow,
         Commodity = hs,
         Weight   = weight,
         Value    = value,
         UV       = uv,
         UV_me    = uv_reporter) %>% 
  sample_n(10) %>% 
    mutate_each(funs(rnd = round(., 2)), 
              starts_with("UV")) 
out_tab %>% 
  printTab(caption = "Example trade flows with outlied unit values")

```

## Imputing using reporter median unit values

Now we correct weight of this outlied trade flows with help of median reporter unit value for a given commodity.

```{r outlier_corrections}
out_tab %>% 
  mutate(Weight_corr = round(Value / UV_me, 0),
         Weight_diff = Weight - Weight_corr) %>%
  select(-Value, -UV, -UV_me) %>% 
  printTab(caption = "Example trade flows with corrected weight")
```

## Imputing using data from trade partner (mirroring)

